Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-06-06 paper2.qmd
section:6; process:write

<!-- 5. So, what are we doing? -->
To address these issues, this study extends Thurstone's general form through a systematic and integrated approach that combines causal and Bayesian inference methods. In addition to potentially enhancing measurement reliability and validity, and improving statistical accuracy in hypothesis testing, this approach offers two key advantages. First, it clarifies the interactions among all actors and processes involved in CJ assessments. Second, it shifts the current comparative data analysis paradigm from passively accepting the BTL model assumptions to actively testing whether those assumptions fit the data under analysis.

@@

<!-- 5. So, what happens when the assumption does not hold? -->
Significant statistical and measurement issues can arise when the assumption of equal dispersions between stimuli does not hold. Specifically, the BTL model may overestimate the trait's reliability, that is, the degree to which the outcome accurately reflects the *true* discriminal differences between stimuli. This overestimation, in turn, results in spurious conclusions about these differences [@McElreath_2020; @Wu_et_al_2022] and, by extension, about the underlying discriminal processes of stimuli. @fig-dispersion also illustrates this scenario when the model's discriminal difference distribution aligns with the thick continuous line for $\sigma_{B}-\sigma_{A}=0$, while the *true* discriminal difference follows any discontinuous line where $\sigma_{B}-\sigma_{A} \neq 0$. Furthermore, if researchers acknowledge that misfit statistics may represent texts with different dispersions or outlying observations, the common CJ practice of excluding stimuli based on these statistics, as seen in @Pollitt_2012a, @Pollitt_2012b, @vanDaal_et_al_2016, and @Goossens_et_al_2018, may unintentionally discard valuable information, introducing bias into the trait estimates [@Zimmerman_1994; @McElreath_2020]. The direction and magnitude of these biases remain unpredictable, as they depend on which stimuli researchers exclude from the analysis.

@@

Finally, similar to what we discussed in the previous section, removing *misfit* judges--that is, judges whose assessments deviate markedly from the shared consensus [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018; @Wu_et_al_2022], and may appear as outliers under the BTL model [@Wu_et_al_2022]--risk discarding valuable information and introducing bias into trait estimates. The direction and magnitude of these biases remain unpredictable, as they depend on which judges researchers exclude from the analysis [@Zimmerman_1994; @OHagan_2018; @McElreath_2020].

@@

<!-- 4. General summary of what we have done. -->
Thus, to address the limitations of Thurstone's Case V and the BTL model, this study extends Thurstone's general form using a systematic, integrated approach that combines causal and Bayesian inference methods. The approach begins by formulating a conceptual model represented by a Structural Causal Model (SCM) and a Directed Acyclic Graph (DAG). This model integrates Thurstone's core theoretical principles, such as the discriminal processes of stimuli, alongside key CJ assessment design features, including judges' bias, sampling procedures, and comparison mechanisms, thereby disentangling the causal processes in the CJ system.

@@

<!-- 5. Are there any additional advantages? -->
Beyond these potential benefits, the approach offers two additional advantages. First, it clarifies the roles and interactions of all actors and processes involved in CJ assessments. Second, it shifts the analytic paradigm from passively accepting the assumptions of the BTL model to actively testing their fit with observed data. Together, these advantages establish a principled framework for evaluating best practices in designing CJ assessments, one that better aligns with the demands of contemporary CJ contexts [@Kelly_et_al_2022], offering new insights into existing research and opening promising avenues for future inquiry.

@@

## Future research directions based on this study's approach {#sec-discussion_RA}
<!-- recording: Sven and Tine 25.06.04; time: 00:05:40 - 00:07:25 -->

Among the many potential directions for future research--both current and new--three avenues deserve particular attention due to their direct impact on the reliability and validity of trait estimates, as well as on the accuracy of statistical inferences. The following sections outline these avenues and explain how our approach provides a rigorous research framework to investigate them.

@@

<!-- 1. The overlooked role of the sampling and comparison mechanisms in CJ -->
Although sampling and comparison mechanisms are central to modern CJ assessments, it is striking that most of the CJ literature has examined them only within a limited scope. Researchers have primarily investigated the effects of adaptive comparative judgment (ACJ) designs on trait reliability [@Pollitt_2012a; @Pollitt_2012b; @Bramley_2015; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021; @Gray_et_al_2024] or proposed practical guidelines for the number of comparisons judges should make [@Verhavert_et_al_2019; @Crompvoets_et_al_2022]. While these studies offer valuable insights, they often overlook the broader role these mechanisms play within the CJ system. Moreover, as this oversight likely stems from a broader lack of conceptual clarity about how these mechanisms function within the system, this study formally incorporates these mechanisms into CJ's conceptual model.

@@

<!-- 4. More complex CJ designs and the implications of MAR or MNAR -->
However, many modern CJ applications rely on more complex assessment designs in which the sampling and comparison mechanisms introduce more intricate forms of missingness--such as *missing at random* (MAR) or *missing not at random* (MNAR) [@Little_et_al_2020]. A prominent example is the previously discussed ACJ designs, where prior judgment outcomes inform the selection of stimulus pairs for subsequent comparisons [@Pollitt_2012a; @Pollitt_2012b; @Bramley_2015]. This outcome-informed pair selection suggests that the comparison mechanism is outcome-dependent, potentially classifying ACJ as a generator of MNAR data. This, in turn, facilitates a more nuanced understanding of the mixed evidence surrounding ACJ's capabilities. While some studies report that the method improves trait reliability [@Pollitt_et_al_2003; @Pollitt_2012a; @Pollitt_2012b], others argue that it artificially inflates these gains [@Bramley_2015; @Bramley_et_al_2019; @Crompvoets_et_al_2020; @Crompvoets_et_al_2022], potentially as a result of outcome-dependent missingness.

@@

<!-- 5. The importance of clarifying assumptions behind CJ assessment designs -->
Nevertheless, regardless of the underlying missingness mechanisms, any CJ assessment design would benefit from explicitly defining its assumptions--a practice supported by our proposed approach. This clarity enables researchers to evaluate how the sampling and comparison mechanisms affect trait estimation and statistical inference within each design. Such assessments are particularly relevant given the common misconception in the CJ literature that Thurstone's model can naturally handle even non-random missing data without compromising the reliability or validity of trait estimates [@Bramley_2008].

@@

<!-- 2. Relevant research question paved by recognizing judges' bias (part II) -->
Moreover, since judges draw on a variety of intricate stimulus characteristics when evaluating complex, multidimensional traits [@vanDaal_et_al_2016; @Lesterhuis_et_al_2018; @Chambers_et_al_2022], and these characteristics account for variation in judgment accuracy [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020; @Gijsen_et_al_2021], it is reasonable to expect that judges' assessments also vary with individual attributes such as gender, age, culture, income, education, training, or expertise [@Kelly_et_al_2022; @Bartholomew_et_al_2020a]. Thus, building on the discussion in @sec-discussion_RA1, researchers could investigate how the selection of judges influences the formation of a "shared consensus" and whether these individual attributes introduce systematic biases or distortions in the observed trait distribution [@Deffner_et_al_2022]. Furthermore, if such attributes indeed undermine the assumption of "sample-freeness," then it becomes essential to explore strategies to control their influence and to determine how many judges--and how many judgments per judge--are needed to produce reliable trait estimates in these contexts. Additionally, it is worth considering whether *repeated measures designs*, in which judges evaluate the same stimulus pairs multiple times [@Lawson_2015], can improve judgment consistency and accuracy.

@@

[LAST THING TO DO]{style="color:blue;"}

<!-- Although the CJ literature clearly defines *misfit judges* as those whose assessments deviate markedly from the "shared consensus" [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018; @Wu_et_al_2022], there has been little discussion about how these misfit judges relate to Thurstonian theory. In particular, it remains unclear which aspects of Thurstone's theory might explain the occurrence of misfit judges. To address this, we propose three statistical perspectives: (1) misfit judges are simply judges who exhibit bias; (2) misfit judges are those whose judgments display greater variability than their peers, and (2) misfit judges are indeed outliers. -->

<!-- Regarding the first perspective, @sec-theory-issue1b explains in detail how, due to the assumption of zero correlation, judges' biases can be miss classified as *misfits*. Meanwhile, @sec-discussion_RA2 discusses how our approach explicitly account for such biases, aiming to reduce the reliance on ad-hoc detection methods--such as misfit statistics--which are poorly suited to the task [@Kelly_et_al_2022]. -->

<!-- Regarding the second perspective -->

<!-- Conversely, regarding the second perspective, the very definition of *misfit judges* leads to their identification as outliers--that is, observations that deviate significantly from the rest of the sample in which they occur [@Grubbs_1969].  -->

<!-- In the previous subsection we argue that ad-hoc detection methods, such as 'misfit' statistics, could be poorly suited for the task of detecting judges bias [@Kelly_et_al_2022]. This hinges in the assumption that the CJ data does have judges biases, but the method does not manages to identify them, leading to false negatives? or lack of sensitivity in the tests. However, there is a related problem which is: these ad-hoc method could be detecting judges' biases that are not there. as in @Wu_et_al_2022  -->

<!-- A similar point can be made for *misfit stimuli*--that is, those that elicit more judgment discrepancies than others [@Pollitt_2004; @Pollitt_2012b; @Pollitt_2012a; @Goossens_et_al_2018]. It remains unclear which aspects of Thurstone's theory might explain the occurrence of misfit stimuli. To address this, we propose two statistical perspectives: (1) misfit stimuli are those whose discriminal processes exhibit greater variability compared to other stimuli, and (2) are statistical outliers. -->

<!-- Regarding the first perspective, @sec-theory-issue1a  -->

<!-- , possibly due to larger discriminal dispersions or because they are genuine outliers--meaning, texts with distinctive characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. In either case, the BTL model neither accounts for these anomalies nor provides tools for addressing them, apart from excluding the "problematic" texts from analysis. -->

<!-- The same could be said for the stimuli. These misfit texts likely arise from two underlying mechanisms. First, a text may exhibit a larger discriminal dispersion, increasing the overlap between its discriminal process and those of multiple other texts. This greater overlap leads to more judgment discrepancies, increasing the likelihood of classifying the text as a misfit. @fig-discriminal_process illustrates this scenario. In the figure, Text B has a wider discriminal dispersion than Text A, enlarging the region where their discriminal processes intersect. This larger overlap represents an area where judges face greater difficulty in determining which text exhibits a higher level of the trait. Second, a text may represent an outlier --one with distinct characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. Such texts are harder to compare and place on the trait scale, resulting in more judgment discrepancies and a misfit classification. In either case, the BTL model is inherently agnostic to these mechanisms and does not provide any means to address them. -->

<!-- Indeed, the presence of the so-called *misfit* texts may already signal these limitations, as these are texts that elicit more judgment discrepancies than others [@Pollitt_2004; @Pollitt_2012b; @Pollitt_2012a; @Goossens_et_al_2018], possibly due to larger discriminal dispersions or because they are genuine outliers--meaning, texts with distinctive characteristics that deviate markedly from the rest of the sample [@Grubbs_1969]. In either case, the BTL model neither accounts for these anomalies nor provides tools for addressing them, apart from excluding the "problematic" texts from analysis. -->

@@

<!-- We advocate for the no removal of judges, but rather identify them and treat the data with the appropriate model, so these 'misfits' do not affect the measurement of the trait -->

<!-- Fortunately, the statistical literature offers a solution to address violations of the equal dispersion assumption in CJ experiments. This solution involves using models that extend traditional approaches to account for the different variability in the stimuli, referred to as over-dispersed models [@McElreath_2020, chap. 12]. -->

@@

@@

@@

<!-- Second, and complementary to the previous point, the statistical literature consistently highlights the risks of using ad hoc procedures to identify "outliers" when the model does not accurately represent the data-generating process. For instance, McElreath (2020, pp. 232) expressed that: "There is a tradition of dropping outliers. People sometimes drop outliers even before a model is fit, based only on standard deviations from the mean outcome value. You should never do that--a point can only be unexpected and highly influential in light of a model. After you fit a model, the picture changes. If there are only a few outliers, and you are sure to report results both with and without them, dropping outliers might be okay. But if there are several outliers, we really need to model them". Similarly, O'Hagan (2018, pp. 366) makes a small example explaining how the average of a parameter--much like a trait score you could get from the BTL model--depends on outliers. Miller (2023), in psychology, explores a similar topic. -->

<!-- You might notice I use the term "outlier" here to describe biased judges. This is because, like statistical outliers, these judges deviate significantly from the rest of the sample in which they occur (Grubbs, 1969). However, the key reason for using this terminology lies in that any ad hoc method for flagging biased judges based on the BTL model--such as misfit statistics--effectively acts as an outlier detection tool. This is due to the fact that the BTL model relies on several strong assumptions, which limit its capacity to capture other features of the data unless supplemented by additional procedures--procedures that, in essence, resemble outlier detection.  -->

<!-- Moreover, from McElreath's text, you might recognize that the practice of removing outliers mirrors the act of discarding biased judges in CJ using misfit statistics, where these statistics are essentially transformations of residual probabilities linked to trait estimates, or as McElreath puts it "standard deviations from the mean outcome values." But what are the potential consequences of wrongfully identifying and discarding misfit judges, you might ask?. Well, this practice risks removing valuable information and introducing bias into trait estimates [Zimmerman, 1994; O'Hagan, 2018; McElreath, 2020]. The direction and magnitude of these biases are unpredictable, depending on which judges are excluded from the analysis. -->

@@

<!-- 1. Challenges for the approach: background knowledge and research context -->
Conversely, this theoretical commitment to causal inference also introduces several practical challenges that applied CJ researchers must navigate. These fall into two main categories: first, acquiring the foundational knowledge necessary to apply the approach effectively; and second, dedicating greater attention to both conceptual and statistical modeling.

>>>>

<!-- 5. So, what are we doing? -->
To address these issues, this study extends Thurstone's general form through a systematic and integrated approach that combines causal and Bayesian inference methods. In addition to potentially enhancing measurement reliability and validity, and improving statistical accuracy in hypothesis testing, this approach offers two key advantages. First, it clarifies the interactions among all actors and processes involved in CJ assessments. Second, it shifts the current comparative data analysis paradigm from passively accepting Case V and the BTL model assumptions to actively testing whether those assumptions fit the data under analysis.

@@

<!-- 5. So, what happens when the assumption does not hold? -->
Significant statistical and measurement issues can arise when the assumption of equal dispersions between stimuli does not hold. Specifically, the BTL model may overestimate the trait's reliability, that is, the degree to which the outcome accurately reflects the *true* discriminal differences between stimuli. This overestimation, in turn, results in spurious conclusions about these differences [@McElreath_2020; @Wu_et_al_2022] and, by extension, about the underlying discriminal processes of stimuli. @fig-dispersion also illustrates this scenario when the model's discriminal difference distribution aligns with the thick continuous line for $\sigma_{B}-\sigma_{A}=0$, while the *true* discriminal difference follows any discontinuous line where $\sigma_{B}-\sigma_{A} \neq 0$. Furthermore, if researchers acknowledge that *misfit statistics* may represent texts with different dispersions or outlying observations, the common CJ practice of excluding stimuli based on these statistics may unintentionally discard valuable information [@Miller_2023], introducing bias into the trait estimates [@Zimmerman_1994; @McElreath_2020]. The direction and magnitude of these biases remain unpredictable, as they depend on which stimuli researchers exclude from the analysis.

@@

Finally, similar to what we discussed in the previous section, removing *misfit* judges--that is, judges whose assessments deviate markedly from the shared consensus [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018; @Wu_et_al_2022], and may appear as outliers under the BTL model [@Wu_et_al_2022]--risk discarding valuable information and introducing bias into trait estimates [@Miller_2023]. The direction and magnitude of these biases remain unpredictable, as they depend on which judges researchers exclude from the analysis [@Zimmerman_1994; @OHagan_2018; @McElreath_2020].

@@

<!-- 4. General summary of what we have done. -->
Thus, to address the limitations of Thurstone's Case V and the BTL model, this study extends Thurstone's general form using a systematic, integrated approach that combines causal and Bayesian inference methods. The approach begins with the development of a conceptual model, represented as a Structural Causal Model (SCM) and a Directed Acyclic Graph (DAG) [@Pearl_2009; @Pearl_et_al_2016; @Gross_et_al_2018; @Neal_2020]. This model integrates Thurstone's core theoretical principles, such as the discriminal processes of stimuli, alongside key CJ assessment design features, including judges' bias, sampling procedures, and comparison mechanisms, thereby disentangling the causal processes underlying the CJ system.

@@

<!-- 5. Are there any additional advantages? -->
Beyond these potential benefits, the approach offers two additional advantages. First, it clarifies the roles and interactions of all actors and processes involved in CJ assessments. Second, it shifts the analytic paradigm from passively accepting the assumptions of Case V and the BTL model to actively testing their fit with observed data. Together, these advantages establish a principled framework for evaluating best practices in designing CJ assessments, one that better aligns with the demands of contemporary CJ contexts [@Kelly_et_al_2022], offering new insights into existing research and opening promising avenues for future inquiry.

@@

## Future research directions using our approach {#sec-discussion_RA}
<!-- recording: Sven and Tine 25.06.04; time: 00:05:40 - 00:07:25 -->

Among the many potential directions for future research--both current and new--three avenues deserve particular attention due to their direct impact on the reliability and validity of CJ trait estimates, as well as on the accuracy of statistical inferences. The following sections outline these avenues and explain how our approach facilitates their investigation.

@@

<!-- 1. The overlooked role of the sampling and comparison mechanisms in CJ -->
Although sampling and comparison mechanisms are central to modern CJ assessments, it is striking that most of the CJ literature has examined them only within a limited scope. Researchers have primarily investigated the effects of adaptive comparative judgment (ACJ) designs on trait reliability [@Pollitt_2012a; @Pollitt_2012b; @Bramley_2015; @Verhavert_et_al_2022; @Mikhailiuk_et_al_2021; @Gray_et_al_2024] or proposed practical guidelines for the number of comparisons judges should make [@Verhavert_et_al_2019; @Crompvoets_et_al_2022]. While these studies offer valuable insights, they often overlook the broader role these mechanisms play within the CJ system. As this oversight likely stems from a broader lack of conceptual clarity about how these mechanisms function within the system, this study decided to formally incorporate these mechanisms into CJ's conceptual model.

@@

<!-- 4. More complex CJ designs and the implications of MAR or MNAR -->
However, many modern CJ applications rely on more complex assessment designs, in which the sampling and comparison mechanisms introduce more intricate forms of missingness — such as *missing at random* (MAR) or *missing not at random* (MNAR) [@Little_et_al_2020]. A prominent example is the previously discussed ACJ designs, where prior judgment outcomes inform the selection of stimulus pairs for subsequent comparisons [@Pollitt_2012a; @Pollitt_2012b; @Bramley_2015]. This pair selection process suggests that ACJ's comparison mechanism is outcome-dependent, potentially classifying the method as a generator of MNAR data. If this classification holds, the mixed findings on ACJ's capabilities become more comprehensible: some studies find that the method improves trait reliability [@Pollitt_et_al_2003; @Pollitt_2012a; @Pollitt_2012b], while others contend that it artificially inflates these gains [@Bramley_2015; @Bramley_et_al_2019; @Crompvoets_et_al_2020; @Crompvoets_et_al_2022].

@@

<!-- 5. The importance of clarifying assumptions behind CJ assessment designs -->
Nevertheless, regardless of the underlying missingness mechanisms, any CJ assessment design would benefit from explicitly defining its assumptions--a practice supported by our approach. This clarity enables researchers to evaluate how the sampling and comparison mechanisms affect trait estimation and statistical inference within each design. Such assessments are particularly relevant given the common misconception in the CJ literature that Thurstone's model can naturally handle even non-random missing data without compromising the reliability or validity of trait estimates [@Bramley_2008].

@@

<!-- 2. Relevant research question paved by recognizing judges' bias (part II) -->
Moreover, since judges draw on a variety of intricate stimulus characteristics when evaluating complex, multidimensional traits [@vanDaal_et_al_2016; @Lesterhuis_et_al_2018; @Chambers_et_al_2022], and these characteristics account for variation in judgment accuracy [@Gill_et_al_2013; @vanDaal_et_al_2017; @vanDaal_2020; @Gijsen_et_al_2021], it is reasonable to expect that judges' assessments also vary with attributes such as gender, age, culture, income, education, training, or expertise [@Kelly_et_al_2022]. Prior studies support this view [@Bartholomew_et_al_2020a; @McMahon_et_al_2015]. Thus, building on the discussion in @sec-discussion_RA1, researchers could further explore how judges' selection influences the formation of a "shared consensus" and whether these attributes introduce systematic biases or distortions in the observed trait distribution [@Deffner_et_al_2022]. Furthermore, if such attributes indeed undermine the assumption of "sample-freeness," researchers should explore ways to control their influence and determine how many judges--and how many judgments per judge--are needed to produce reliable trait estimates in these contexts. Additionally, it is worth considering whether *repeated measures designs*, in which judges evaluate the same stimulus pairs multiple times [@Lawson_2015], can improve judgment consistency and accuracy. Our approach provides the necessary structure to rigorously investigate these questions.

@@

<!-- without disentangling all the potential processes the concept of misfits is strange, because it is not defined from Thustonian theory. As it is how is treated in CJ literature, misfit are outliers. People have theories on what makes misfits, but the approach now allows to test some hypothesis that might generate outliers under the BTL model, but no longer considered outliers in our approach  -->

<!-- the approach makes it possible to have hypothesis about why judges are labeled as misfits.  -->

Although the CJ literature clearly defines *misfit* judges and stimuli, researchers have rarely examined how these observations relate to Thurstonian theory. In particular, they have not identified which elements of Thurstone's theory might explain the occurrence of misfits. This disconnect likely stems from the fact that CJ researchers derive misfit statistics from methods such as residual analysis and outlier detection rather than from Thurstonian principles [@Pollitt_2012a; @Pollitt_2012b; @Wu_et_al_2022]. They typically define *misfit judges* as those whose assessments diverge significantly from the "shared consensus" [@Pollitt_2012a; @Pollitt_2012b; @vanDaal_et_al_2016; @Goossens_et_al_2018; @Wu_et_al_2022], and *misfit stimuli* as those that elicit more judgment discrepancies than others [@Pollitt_2004; @Pollitt_2012a; @Pollitt_2012b; @Goossens_et_al_2018]. Both definitions closely mirror the statistical concept of outliers--that is, observations that deviate markedly from the rest of the sample in which they occur [@Grubbs_1969]. However, this resemblance extends beyond these definitions. CJ researchers also identify misfits using standard outlier detection procedures, such as transforming BTL model residuals into statistics of interest and comparing them against a predefined threshold.

@@

[After identifying misfit observations, researchers often propose various hypotheses to explain their presence. Yet, without providing evidence for any of these explanations, they typically exclude these observations from analysis. This exclusion process raises three major concerns: one related to the reliability of the methods used to identify misfits, a second concerning the potential impact of their removal on trait estimates and inferences, and a third regarding the hypotheses used to explain the presence and exclusion of misfits.]{style="color:blue;"}

[First, regarding the reliability of identification methods, it is relevant to note that outlying observations can only be defined relative to a specific model [@McElreath_2020]. Consequently, methods based on models such as the BTL--which rest on strong assumptions--should be treated with caution, as these assumptions may prevent the model from accurately capturing the underlying data-generating process of the CJ system, thereby potentially compromising the identification of misfits.]{style="color:blue;"}

@@

[Second, concerning the impact of removing misfits, the statistical literature warns that ad-hoc outlier exclusion procedures risk discarding valuable information [@Miller_2023] and introducing bias into trait estimates. The direction and magnitude of this bias are often unpredictable, as they depend on which observations researchers exclude from the analysis [@Zimmerman_1994; @OHagan_2018; @McElreath_2020].]{style="color:blue;"}

@@

[Lastly, regarding the hypotheses explaining the presence and exclusion of misfits, it is important to note that a significant number of these hypotheses could not be tested within the BTL model due to its strong assumptions and limited flexibility. This is the reason why there is a lack of supporting evidence for misfit observations. In contrast, our approach enables researchers to test these hypotheses. For instance, researchers could examine whether misfit judges are simply those who exhibit (an outlying degree of) systematic bias or whether they display greater variability in their judgments compared to their peers. Similarly, in the case of misfit stimuli, researchers could investigate whether their status reflects an increased variability in their discriminal processes or stems from their classification as outlying stimuli.]{style="color:blue;"}

[Moreover, since outliers are not necessarily "bad data" [@McElreath_2020], our approach offers a principled alternative that retains misfits in the analysis without compromising trait estimation. This alternative involves adapting our proposed models into more sophisticated, robust measurement models [@Wainer_et_al_1978]--a broad class of procedures designed to reduce the sensitivity of parameter estimates to mild or moderate departures from model assumptions [@Everitt_et_al_2010].]{style="color:blue;"}

@@

@@

<!-- 1. Challenges for the approach: background knowledge and research context -->
Moreover, this theoretical commitment to causal inference also introduces several practical challenges that applied CJ researchers must navigate. These fall into two main categories: first, acquiring the foundational knowledge necessary to apply the approach effectively; and second, dedicating greater attention to both conceptual and statistical modeling.
